### 2024-12-04 manuscript.tex
Update on Overleaf.

ERA5 \citep{era5} is a reanalysis produced by the ECMWF.  It is based on a
numerical weather prediction model IFS version CY41R2.  The horizontal
resolution is 0.25° in latitude and longitude, which is about 28 km in the
North--South direction and 14 km in the East--West direction at 60°S.
Internally, the model uses 137 vertical levels. Here, we use output at 1-hourly
instantaneous time intervals, except for radiation quantities, which are
accumulations (from these we calculate daily means).  Vertically resolved
quantities are made available on 37 pressure levels.

@@

The ANN achieved 65\% sensitivity and 87\% specificity when the true positive
rate (26\%) was made to match observations. The receiver operating
characteristic curve is shown in Fig. \ref{fig:ann}c. We considered these rates
satisfactory for the purpose of filtering precipitation profiles. Fig.
\ref{fig:ann}d shows examples of the predicted precipitation compared to
human-performed observations.

@@

We partitioned our data into two mutually exclusive subsets by cyclonic
activity. For this purpose, we used a cyclone tracking algorithm to identify
extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis
and ICON data. We used the open source cyclone tracking package CyTRACK
\citep{perez-alarcon2024}.  Generally, what constitutes an EC is considered
relatively arbitrary due to the very large variability of ECs \citep{neu2013}.
We used the mean sea level pressure field and horizontal wind speed fields as
input to the CyTRACK algorithm. The algorithm uses pressure and wind speed
thresholds as well as tracking across time steps to identify cyclone centres
and radii. With this information, we could classify geographical areas as
either cyclonic or non-cyclonic. Due to a relatively small total area covered,
we chose a circle of a double radius (relative to one identified by CyTRACK)
centred at the cyclone centre as a cyclonic area for every time step and
cyclone. All other areas were identified as non-cyclonic. For identifying
cyclones in the observations and the reanalyses, ERA5 pressure and wind fields
were used as the input to CyTRACK.  This is justified by the fact that the
large-scale pressure and wind fields in ERA5 are likely sufficiently close to
reality. For identifying cyclones in ICON, its own pressure and wind fields
were used as the input to CyTRACK, because the model is free-running, and thus
the pressure and wind fields are different from reality.

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relevant for the subsequent analysis, because these
conditions are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a, b show a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied to the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, there are clear differences particularly in the highest occurence rate regions -- over the Amundsen Sea, which is
more cyclonic in the reanalysis, and around Cape Adare, which is more cyclonic
in ICON. These differences might, however, stem from the relatively short time
periods of comparison (4 years) and the fact that the model is free-running.

Fig. \ref{fig:cyclone-stability}c, d show a geographical distribution of the
relatively stable and unstable conditions as determined by the LTS (Section
\ref{sec:cyclone-stability}). Relatively unstable conditions are prevalent
in the middle SO (50--65°S), which might be explained by the relatively cold
near-surface air overlying the relatively warm sea surface. Relatively
stable conditions are prevalent elsewhere over the SO. The distribution is also
less zonally symmetric than the cyclonic activity.  In the high-latitude SO,
the presence of sea ice might have substantial stabilising effect
\citep{knight2024}. The ERA5 reanalysis is also substantially more stable than
ICON across the whole region.

@@

We analysed about 2300 radiosonde profiles south of 40°S from the 24 RV
\emph{Polarstern} voyages, MARCUS, NBP1704, TAN1702, and TAN1802. Spatially and
temporally colocated profiles were taken from ICON and the reanalyses.
Because the time period covered by the ICON model output (2021--2024) was
different from the time period covered by the observations (2010--2021), when
comparing with the model, we first had to remap the observation time to model
time by taking the same time relative to the start of the year. Consequently,
we also had four virtual/model profiles (one for each year of 2021--2024) for
each observed profile.
The profiles were partitioned into the same subsets as
above (Sections \ref{sec:cloud-occurrence} and \ref{sec:cloud-cover}).  Apart
from relative humidity, we focus on comparing virtual potential temperature
($\theta_v$) due to its role in low-level tropospheric stability, being one of
the primary factors affecting shallow convection and the associated low-level
cloud formation and dissipation. The observed and model profiles of virtual
potential temperature are shown in Fig. \ref{fig:potential-temperature}.

@@

The time period of ICON is relatively short, with only four full years of
simulation available. Moreover, the simulation is free-running, which means
that observations had to be temporally mapped to this time period (at the same
time relative to the start of the year) for the comparison. For these reasons,
one can expect the results to be slightly different due to reasons unrelated to
model biases, such as different weather conditions and the phase of climate
oscillations such as the ENSO in the observations and the model.

>>>>

ERA5 \citep{era5} is a reanalysis produced by the ECMWF.  It is based on a
numerical weather prediction model IFS version CY41R2.  It uses the \cite{tiedtke1993} prognostic cloud scheme and \cite{forbes2014} for mixed-phase clouds. The horizontal
resolution is 0.25° in latitude and longitude, which is about 28 km in the
North--South direction and 14 km in the East--West direction at 60°S.
Internally, the model uses 137 vertical levels. Here, we use output at 1-hourly
instantaneous time intervals, except for radiation quantities, which are
accumulations (from these we calculate daily means).  Vertically resolved
quantities are made available on 37 pressure levels.

@@

The ANN achieved 65\% sensitivity and 87\% specificity when the true positive
rate (26\%) was made to match observations. The receiver operating
characteristic curve is shown in Fig. \ref{fig:ann}c. We considered these rates
satisfactory for the purpose of filtering precipitation profiles. Fig.
\ref{fig:ann}d shows examples of the predicted precipitation compared to
human-performed observations. The main ANN (`ANN` in Fig. \ref{fig:ann}) was trained on all data, and ancillary ANNs (`ANN2` in Fig \ref{fig:ann}) were trained on all data except for each voyage. These were used for testing the results relative to each voyage.

@@

We partitioned our data into two mutually exclusive subsets by cyclonic
activity. For this purpose, we used a cyclone tracking algorithm to identify
extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis
and ICON data. We used the open source cyclone tracking package CyTRACK
\citep{perez-alarcon2024}.  Generally, what constitutes an EC is considered
relatively arbitrary due to the very large variability of ECs \citep{neu2013}.
We used the mean sea level pressure field and horizontal wind speed fields as
input to the CyTRACK algorithm. The algorithm uses pressure and wind speed
thresholds as well as tracking across time steps to identify cyclone centres
and radii in each time step. With this information, we could classify geographical areas as
either cyclonic or non-cyclonic. Due to a relatively small total area covered by cyclones (as identified by the cyclone centre and radius),
we chose a circle of a double radius (relative to one identified by CyTRACK)
centred at the cyclone centre as a cyclonic area for every time step and
cyclone. All other areas were identified as non-cyclonic. For identifying
cyclones in the observations and the reanalyses, ERA5 pressure and wind fields
were used as the input to CyTRACK.  This is justified by the fact that the
large-scale pressure and wind fields in ERA5 are likely sufficiently close to
reality. For identifying cyclones in ICON, its own pressure and wind fields
were used as the input to CyTRACK, because the model is free-running, and thus
the pressure and wind fields are different from reality.

@@

Fig. \ref{fig:cyclone-stability}a, b show a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied to the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, there are clear differences particularly in the highest occurrence rate regions -- over the Amundsen Sea, where the Amundsen Sea low is located \cite{turner2013} and which is
more cyclonic in the reanalysis, and around Cape Adare, which is more cyclonic
in ICON. These differences might, however, stem from the relatively short time
periods of comparison (4 years) and the fact that the model is free-running.

Fig. \ref{fig:cyclone-stability}c, d show a geographical distribution of the
relatively stable and unstable conditions as determined by the LTS (Section
\ref{sec:cyclone-stability}). Relatively unstable conditions are prevalent
in the mid-to-high SO (with respect to our SO partitioning; 50--65°S), which might be explained by the relatively cold
near-surface air overlying the relatively warm sea surface. Relatively
stable conditions are prevalent elsewhere over the SO. The distribution is also
less zonally symmetric than the cyclonic activity.  In the high-latitude SO,
the presence of sea ice might have substantial stabilising effect
\citep{knight2024}. The ERA5 reanalysis is also substantially more stable than
ICON across the whole region.

@@

In order to link the cloud biases to the local physical conditions, we analysed about 2300 radiosonde profiles south of 40°S from the 24 RV
\emph{Polarstern} voyages, MARCUS, NBP1704, TAN1702, and TAN1802. Spatially and
temporally colocated profiles were taken from ICON and the reanalyses.
Because the time period covered by the ICON model output (2021--2024) was
different from the time period covered by the observations (2010--2021), when
comparing with the model, we first had to remap the observation time to model
time by taking the same time relative to the start of the year. Consequently,
we also had four virtual/model profiles (one for each year of 2021--2024) for
each observed profile.
The profiles were partitioned into the same subsets as
above (Sections \ref{sec:cloud-occurrence} and \ref{sec:cloud-cover}).  Apart
from relative humidity, we focus on comparing virtual potential temperature
($\theta_v$) due to its role in low-level tropospheric stability, being one of
the primary factors affecting shallow convection and the associated low-level
cloud formation and dissipation. The observed and model profiles of virtual
potential temperature are shown in Fig. \ref{fig:potential-temperature}.

@@

The time period of ICON is relatively short, with only four full years of
simulation available. Moreover, the simulation is free-running, which means
that observations had to be temporally mapped to this time period (at the same
time relative to the start of the year) for the comparison. For these reasons,
one can expect the results to be slightly different due to reasons unrelated to
model biases, such as different weather conditions, partially accounted for by the cyclone and stability subsetting, and the phase of climate
oscillations such as the ENSO in the observations and the model.
